---
title: "Introduction to ggtranscript"
author: 
  - name: David Zhang
    affiliation:
    - UCL
    email: dyzhang32@gmail.com
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('ggtranscript')`"
vignette: >
  %\VignetteIndexEntry{Introduction to ggtranscript}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# Basics

To be completed...

## Install `ggtranscript`

`R` is an open-source statistical environment which can be easily modified to enhance its functionality via packages. `r Biocpkg("ggtranscript")` is a `R` package available via the [Bioconductor](http://bioconductor.org) repository for packages. `R` can be installed on any operating system from [CRAN](https://cran.r-project.org/) after which you can install `r Biocpkg("ggtranscript")` by using the following commands in your `R` session:

```{r "install", eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("ggtranscript")

## Check that you have a valid Bioconductor installation
BiocManager::valid()
```
